Objectification in Action: Self- and Other-Objectification in Same-gender Interactions

### 2020-09-25 Final_Manuscript.tex
updated SEM figure

\abstract{
Empirical evidence has only found links between objectification,
self-objectification, and negative outcomes for women within
interpersonal interactions between male-female pairs. The purpose of the
present study was to extend past research and consider the relationships
between objectification and self-objectification within interactions
between female pairs. Women were brought into the laboratory and
interacted in same-sex dyads. Dyadic analysis was utilized to detect
whether partners' objectification of each other affected state
self-objectification, as well as the resulting feelings of comfort and
authenticity during the interaction. After the interaction, participants
completed a questionnaire measuring, among other things, cognitive
performance, career aspirations, and relationship agency. Results
revealed a significant positive effect of being objectified by a female
partner on women's self-objectification. There was no significant
relationship between self-objectification and authenticity. Further,
although there were significantly negative effects on career aspirations
and relationship agency resulting from a lack of relationship
authenticity, there was no evidence that this is due to feelings of
sexual objectification. The significant partner effect of
objectification on actor self-objectification suggests that women being
objectified by other women may result in feelings of
self-objectification, and such research has powerful implications for
the ways that women interact in both sexual and non-sexual settings.

@@

Roberts and Fredrickson (1997) Objectification Theory, suggests that, in
addition to being steeped in a culture that objectifies women, women are
objectified in actual interpersonal encounters. The negative effects of
this interpersonal objectification for women are theorized be the
strongest when it is a man doing the objectifying (Calogero, 2004; Gay
\& Castano, 2010; Gervais, Vescio, \& Allen, 2011), but what are the
consequences when a woman objectifies another woman in an interaction?
Psychological researchers studying the sexual objectification of women,
have in recent years explored the interpersonal process of the
objectification, finding evidence for self-objectification in the target
as a proximal consequence of being objectified (Garcia, Earnshaw, \&
Quinn, 2016, Strelan and Pagoudis (2018), Riemer, Sáez, Brock, and
Gervais (2020)). In a recent review, Gervais, Sáez, Riemer, and Klein
(2020) organized the extant literature on interpersonal
self-objectification and proposed a theoretical model called the Social
Interaction Model of Objectification (SIMO), however this model focuses
on mixed-gender encounters. Although there is evidence that women
objectify other women (Loughnan et al., 2015a; Puvia \& Vaes, 2013),
there are currently no studies investigating the process of
interpersonal objectification in interactions between women. Is
objectification by a female interaction partner related to an increase
in self-objectification for the woman being objectified? Further, does
objectification by a woman have the same downstream negative
consequences for women as being objectified by a male interaction
partner? The current study begins to address these questions by
investigating interpersonal objectification among interacting pairs
where both partners identify as women.

@@

Self-objectification is a psychological process that provides a
cognitive mechanism to translate the experiences of interpersonal
objectification (Loughnan et al., 2015a) to negative mental health
outcomes (e.g., anxiety, self-esteem, and cognitive performance)
(Calogero, Tantleff-Dunn, \& Thompson, 2011; Moradi \& Huang, 2008).
Calogero et al. (2011) proposes that self-objectification can be
conceptualized as a learned trait, or \emph{trait} self-objectification
(TSO). Furthermore, self-objectification can also be elicited
momentarily, for example, when viewing sexualized images in movies and
magazines (Morry \& Staska, 2001), when trying on sexualizing clothing
(B. L. Fredrickson, Roberts, Noll, Quinn, \& Twenge, 1998), or during
video chat when you believe your interaction partner can only see your
body (not face; Saguy, Quinn, F Dovidio, \& Pratto, 2010). This
momentary self-objectification is referred to as \emph{state}
self-objectification (SSO; Bernard, Legrand, \& Klein, 2018; Calogero et
al., 2011; Moradi \& Huang, 2008) and is characterized by feeling like a
body rather than a full self within a particular moment, instance, or
context (Roberts \& Fredrickson, 1997).

There is evidence that being objectified by another person during an
interaction elicits SSO. For example, in a face-to-face mixed-gender
dyadic study, Garcia et al. (2016) found that men's reported
objectification of their female interaction partner was associated with
increased self-objectification as reported by their female partner.
Further, this negative effect of the men's objectification on women's
SSO was strongest for women higher in TSO. The current study uses this
same dyadic paradigm to study same-gender, woman-woman interacting
pairs.

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), often times feel \enquote{looked at}
within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and are likely to internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Interpersonal objectification has been found to have negative
consequences for women. Gervais et al. (2011) found that an objectifying
gaze by a male interaction partner (confederate) was associated with
lower math performance for women, than being objectified by a female
interaction partner. Moreover, perhaps the most adverse effect of
interpersonal objectificaiton is that being objectified socializes girls
and women to treat themselves as objects to be looked at and evaluated
(Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972; B. L. Fredrickson et
al., 1998). Gervais et al. (2020) recently reviewed the research on
interpersonal objectification and organized our current theoretical
understanding of this process in the Social Interaction Model of
Objectification (SIMO). What is clear from Gervais et al. (2020)'s
review is that we know little about women-on-women interpersonal
objectification. Could there also be negative consequences of this type
of interpersonal objectification?

Interpersonal objectification can also occur within the context of
romantic relationships, in addition to interactions with strangers.
Among heterosexual male and female college students,
self-objectification has been found to be positively associated with the
extent to which they objectify their romantic partners (Zurbriggen,
Ramsey, \& Jaworski, 2011). Researchers have also studied interpersonal
objectification and self-objectification in romantic relationships
(Strelan \& Pagoudis, 2018). Evidence is mixed, suggesting that women
can sometimes have positive outcomes from appearance valuation within
the context of an established romantic relationship. There is also
evidence that they experience less SSO after appearance comments within
romantic relationships (Meltzer, 2020), but less is known about first
dates. An important caveat to this literature is that it is only about
heterosexual relationships. A recent study has investigated the effect
of self-objectification on the reduction of relationship building skills
in general (including same-sex friendships) (Terán, Jiao, \& Aubrey,
2020). Both romantic relationships and stranger encounters have been
studied in the context of objectification, but neither of these
literatures have investigated specifically same-gender relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
for women would suggest no (Gay \& Castano, 2010; Gervais et al., 2011;
Saguy et al., 2010, Calogero (2004); Yilmaz \& Bozo, 2019), but, there
is evidence that women do objectify other women (Harsey \& Zurbriggen,
2020; Loughnan et al., 2015a; Puvia \& Vaes, 2013) and to would be
helpful to understand the consequencs of this intragroup objectification
for women. When women objectify other women, does it lead to
self-objectification in the same way men's objectification of women does
(Garcia et al., 2016)? If so, does the self-objectification experienced
in these same-gender interactions have the same negative consequences
for authenticity in that interaction?

@@

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected that TSO
would moderate this relationship, amplifying the positive assocaition
between being objectified and SSO for women higher in TSO. We
hypothesized that SSO would, in turn, lead to feelings of inauthenticity
(although there is ample support against this prediction from the
identity threat within intergroup interactions literature; Rollero,
2016). Further, we hypothesized that these feelings of inauthencity
would then be associated with reduced feelings of agency in romantic
relationships, reduced career aspiration, and reduced cognitive
performance. In summary, we expected to find a positive relationship
between other-objectification by one's partner and state
self-objectification. We also expected to find a negative relationship
between self-state objectification and interaction authenticity, and
that interaction authenticity will be positively related to cognitive
performance, relationship agency, and career aspirations.

These hypotheses are considered exploratory, given the lack of prior
research theorizing about these linkages.

Here we would like to note that Puvia and Vaes (2013) would
alternatively predict that women's tendency to self-objectify (TSO)
leads them to objectify other women (more precisely, to dehumanize a
sexualized woman). In addition, Strelan and Hargreaves (2005) and Harsey
and Zurbriggen (2020) would predict the TSO to other-objectification
link, further theorizing that this relationships is mediated by the
woman's own state self-objectification (SSO). This limited, but extant,
set of literature on woman-woman objectification points to a possible
alternative model from the model hypothesized in the current study.
Where appropriate, we report results considering the causal directions
implied by this alternative mediation model.

@@

Data from two demographically similar higher education institutions in
the Northeast United States were combined to create the final analysis
sample (\emph{N =} 64) used in this study. In the measures section that
follows we refer to them as Sample 1 and Sample 2. Thirty-two previously
unacquainted self-identifying female dyads (64 total participants) from
institutions in the Northeast of the United States participated in this
study. Sample 1 (\emph{N =} 24) is from a co-ed liberal arts college and
Sample 2 (\emph{N =} 40) is from a women's liberal arts college. More
specifically, twelve of the pairs were students at a co-ed liberal arts
college, while the remaining twenty pairs attended a women's liberal
arts college. {[}Footnote: Initially, data was collected from same-sex
and mixed-sex dyads that comprised of male and female gendered
individuals. Sample 1 originally consisted of twenty-two pairs, twelve
men and thirty-two women. Twenty-three pairs made up of forty-three
women and one man, as well as two participants who did not identify with
either gender category, formed Sample 2. For consistency, we limited
participant data to same sex female pairs at the two colleges.{]} All
results presented below are from models including sample as a control
variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were cross-racial minority group pairs. 64.06\%
of the sample identified as heterosexual, and 25\% identified as gay,
lesbian or bisexual.

@@

To conceptualize participants' career aspirations after the interaction,
we used the 10-item adaptation of P. Gray and M. OBrien (2007)'s Career
Aspiration Scale employed in Garcia et al. (2016), which asked
participants to consider how true 10 statements were in regard to their
future careers on a scale from 0 (not at all true of me) to 4 (very true
of me). Items include \enquote{I hope to become a leader in my career
field} and \enquote{I hope to move up through any organization or
business I work in.} Items were fairly reliable (\(\alpha\) = 0.73 for
Sample 1, \(\alpha\) = 0.80 for Sample 2).

@@

Trait self-objectification (TSO) was assessed using the
Self-Objectification Questionnaire (B. L. Fredrickson et al., 1998; M.
Noll \& L. Fredrickson, 1998), which evaluates the extent to which
individuals view their bodies in observable versus non-observable ways.
The questionnaire asked participants to rank order both appearance and
functional aspects of their bodies, from 1 (least important) to 10 (most
important), with respect to physical self-concepts. Of the ten body
attributes, five of the items were appearance-based (weight, sex appeal,
physical attractiveness, firm/sculpted muscles and body measurements),
and five of the items were competence-based (strength, physical
coordination, energy level, health and physical fitness). Difference
scores were computed by subtracting the sum of the 5 functional
aspects/competence attributes (e.g., health, strength) from the sum of
the 5 physical self-concepts/appearance attributes (e.g., physical
attractiveness, weight), and all measures were multiplied by -1, as was
done in Garcia et al. (2016), so that positive scores indicated greater
TSO.

@@

@@

\begin{figure}
\includegraphics[width=400px]{theor_figure} \caption{Theoretical Model from Garcia et al. (2016).}\label{fig:theorfigure}
\end{figure}

This study sought to detect whether partners' objectification of one
another affected state self-objectification (SSO). Specifically, we were
interested in testing the relationship between state-other
objectification and SSO, and how SSO in turn, affects feelings of
inauthenticity during the interaction. In addition, we also tested if
the effect of other-objectification in an interaction on SSO is only
present for those women who are high in trait self-objectification.
Further, we investigated the relationships between experiencing
interaction inauthenticity and relationship agency, career aspirations,
and cognitive performance.

While past studies investigating objectification in interactions used
dyadic path analysis (Garcia et al., 2016), the current study used
multilevel modeling proceedures. Dyadic analyses for distinguishable
dyads (e.g., mixed-gender interacting pairs) is more natural in
Structural Equation Modeling (SEM) than it is for indistinguishable
dyads (e.g., same-gender interacting pairs), as we have here (Garcia,
Kenny, \& Ledermann, 2015; Ledermann \& Kenny, 2017). One reason for
this asymmetry is that, due to the arbitrary distinctions made between
\enquote{partner 1} and \enquote{partner 2} in indistinguishable dyads,
many estimates need to be fixed to be equal (i.e., paths, variances,
covariances, endogenous intercepts, and exogenous means), but these
equality constraints should not then be considered in the degrees of
freedom calculations for fit estimations (Olsen \& Kenny, 2006).
Further, Olsen and Kenny (2006) details how a new independence model and
the corresponding fit measure should be re-calculated for
indistinguishable dyads models. The current study uses dyadic multilevel
modeling (MLM) to test all relationships and mediation patterns. The
online supplementary materials contains analysis using SEM. See
Ledermann and Kenny (2017) for a more complete discussion of the
considerations for using SEM versus MLM for dyadic analysis.

Testing our hypotheses in the current, same-gender, sample, involves
using the Actor-Partner Independence Model (APIM) approach for each
outcome variable (i.e., endogenous variable in
Figure~\ref{fig:theorfigure}). Thus, we ran five APIM's to test all the
hypothesized relationships. See Figure~\ref{fig:apim} for a basic APIM
model. The APIM includes effects due to one's own, as well as one's
partner's, predictor variables (\(X\)'s) on the one's own outcome
variable (\(Y\)). The current study deals with indistinguishable dyads,
meaning the designation of who is designated as \enquote{actor} and who
is designated as \enquote{partner} is arbitrary. These analyses are
considered exploratory, given the lack of prior research theorizing
about these linkages.

@@

\begin{figure}
\includegraphics[width=400px]{SEMfigure} \caption{Path Analysis Model with Estimates}\label{fig:semfigure}
\end{figure}

All model estimates and p-values are found in Table~\ref{tab:mlm} and
the relationships with estimates included are depicted in
Figure~\ref{fig:semfigure}.

@@

As is the case in all correlational studies, we cannot be sure about the
causal direction. The model used by the current study asserts that being
objectified by a woman causes women to self-objectify, but it could be
that women's SSO causes them to be objectified by their interaction
partner. This latter interpretation is theoretically interesting given
that, in both the Garcia et al. (2016) study and in the current study,
it is \emph{women's} state self-objectification that relates to being
objectified by one's partner. Objectification Theory (Roberts \&
Fredrickson, 1997), as well as some past experimental studies (Saguy et
al., 2010) suggest that at least one direction of the causal flow is
from other-objectification to SSO. Other studies, especailly those
investigating explicitly women objectifying other women (Puvia \& Vaes,
2013), have relately found evidence that women's \emph{trait}
self-objectification causes women to objectify (dehumanize) other women
and this link is mediated by \emph{state} self-objectification. It could
follow that we tend to objectify other women who objectify us and
through a process of TSO causes SSO which in turn causes objectifation
of one's partner, which in turn causes one's partner to objectify us.
The current study does not provide any evidence of this alternative
model, however---TSO was not significantly correlated with SSO and
actor's objectification was not correlated with partner's
objectification.

>>>>

\abstract{
Extant research on interpersonal objectification has focused mostly on
links between objectification, self-objectification, and negative
outcomes for women within mixed-gender interactions (Garcia, Earnshaw,
\& Quinn, 2016; Gervais, Sáez, Riemer, \& Klein, 2020). The purpose of
the present study was to extend past research on interpersonal
objectification to interactions between pairs of women. Women were
brought into the laboratory and interacted face-to-face in same-sex
dyads. Dyadic analysis (Kenny, Kashy, \& Cook, 2006) was utilized to
detect whether partners' objectification of each other affected state
self-objectification, as well as the resulting feelings of comfort and
authenticity during the interaction. Results revealed a significant
positive relationship between being objectified by a woman interaction
partner and women's own self-objectification (a partner effect). There
was no significant relationship between self-objectification and
interaction inauthenticity. Further, there were significant negative
effects of inauthenticity on career aspirations and relationship agency
(actor effects). The significant partner effect of objectification on
self-objectification suggests that women being objectified by other
women may result in feelings of self-objectification, and this finding
opens up questions for future research about how interpersonal
objectification might function in situations without an immediate male
gaze.

@@

Barbara L Fredrickson and Roberts (1997) Objectification Theory,
suggests that, in addition to being steeped in a culture that
objectifies women, women are objectified in actual interpersonal
encounters. The negative effects of this interpersonal objectification
for women are theorized be the strongest when it is a man doing the
objectifying (Calogero, 2004; Gay \& Castano, 2010; Gervais, Vescio, \&
Allen, 2011), but what are the consequences when a woman objectifies
another woman in an interaction? Psychological researchers studying the
sexual objectification of women, have in recent years explored the
interpersonal process of the objectification, finding evidence for
\emph{self}-objectification in the target as a proximal consequence of
being objectified (Garcia et al., 2016; Riemer, Sáez, Brock, \& Gervais,
2020; Strelan \& Pagoudis, 2018). In a recent review, Gervais et al.
(2020) organized the extant literature on interpersonal
self-objectification and proposed a theoretical model called the Social
Interaction Model of Objectification (SIMO). The SIMO is useful for
studying mixed-gender interactions, but what about interactions among
women? Although there is evidence that women objectify other women
(Loughnan et al., 2015a; Puvia \& Vaes, 2013), studies investigating the
process of interpersonal objectification in interactions between women
is scarce.

Is objectification by a female interaction partner related to an
increases in self-objectification for the woman being objectified?
Further, does objectification by a woman have the same downstream
negative consequences for women as being objectified by a male
interaction partner? The current study begins to address these questions
by investigating interpersonal objectification among interacting pairs
where both partners identify as women.

@@

Self-objectification is a psychological process that provides that
translates the experiences of interpersonal objectification {[}Loughnan
et al. (2015a); Gervais et al. (2020) into negative mental health
outcomes (e.g., interaction inauthenticity, anxiety, self-esteem, and
cognitive performance) (Calogero, Tantleff-Dunn, \& Thompson, 2011;
Moradi \& Huang, 2008). Calogero et al. (2011) proposes that
self-objectification can be conceptualized as a learned trait, or
\emph{trait} self-objectification (TSO), whereby one abopts a habitual
third-person perspective on one's own appearance. Furthermore,
self-objectification can also be elicited momentarily, for example, when
viewing sexualized images in movies and magazines (Morry \& Staska,
2001), when trying on sexualizing clothing (Barbara L Fredrickson,
Roberts, Noll, Quinn, \& Twenge, 1998), or during video chat when you
believe your interaction partner can only see your body (not face;
Saguy, Quinn, F Dovidio, \& Pratto, 2010). This momentary
self-objectification is referred to as \emph{state} self-objectification
(SSO; Bernard, Legrand, \& Klein, 2018; Calogero et al., 2011; Moradi \&
Huang, 2008) and is characterized by feeling like a body rather than a
full self within a particular moment, instance, or context.

There is evidence that being objectified by another person during an
interaction elicits SSO. For example, Loughnan, Baldissarri, Spaccatini,
and Elder (2017) found that for women, imagining a time when they were
objectified by another person cause reductions in human traits
attributed to the self, and further, the gender of the person doing the
objectification was not a moderating factor. In addition, in a
face-to-face mixed-gender dyadic study, Garcia et al. (2016) found that
men's reported objectification of their female interaction partner was
associated with increased self-objectification as reported by their
female partner. Further, this negative effect of the men's
objectification on women's SSO was strongest for women higher in TSO.
The current study uses this same dyadic paradigm to study same-gender,
woman-woman interacting pairs.

@@

Studies have shown that within social encounters women are gazed at more
than men (Briton \& Hall, 1995), often times feel \enquote{looked at}
within interpersonal interactions (Argyle \& Williams, 1969;
Miles-McLean et al., 2015), and are likely to internalize the
objectifying gaze on their physical self (Puvia \& Vaes, 2013; Young,
1979). Interpersonal objectification has been found to have negative
consequences for women. Gervais et al. (2011) found that an objectifying
gaze by a male interaction partner (confederate) was associated with
lower math performance for women, than being objectified by a female
interaction partner. Perhaps the most adverse negative consequence of
interpersonal objectificaiton is that being objectified socializes girls
and women to routinely treat themselves as objects to be looked at and
evaluated (Bartky, 1990; Berger, Cohen, \& Zelditch Jr, 1972; Barbara L
Fredrickson et al., 1998). Gervais et al. (2020) recently reviewed the
research on interpersonal objectification and organized our current
theoretical understanding of this process in the Social Interaction
Model of Objectification (SIMO). What is clear from Gervais et al.
(2020)'s review is that we know little about women-on-women
interpersonal objectification. Could there also be negative consequences
for women of interpersonal objectification by women?

Researchers have also studied interpersonal objectification and
self-objectification in romantic relationships (Strelan \& Pagoudis,
2018), in addition to interactions with strangers. Among heterosexual
male and female college students, self-objectification has been found to
be positively associated with the extent to which they objectify their
romantic partners (Zurbriggen, Ramsey, \& Jaworski, 2011). Evidence is
mixed, suggesting that women can sometimes have positive outcomes from
appearance valuation within the context of an established romantic
relationship. There is also evidence that they experience less SSO after
appearance comments within romantic relationships (Meltzer, 2020), but
less is known about first dates. An important caveat to this literature
is that it is only about heterosexual relationships. A recent study has
investigated the effect of self-objectification on the reduction of
relationship building skills in general (including same-sex friendships)
(Terán, Jiao, \& Aubrey, 2020). Both romantic relationships and stranger
encounters have been studied in the context of objectification, but
neither of these literatures have investigated specifically same-gender
relationships.

There is evidence that the experience of state self-objectification in
mixed-gender contexts (stranger and romantic) has negative consequences
for women, but what about in the context of interactions with other
women? Is the self-objectification experienced in mixed-gender
interactions associated with the same negative process as that
experienced in same-gender interactions? The ample research
demonstrating that the male gaze has a particularly detrimental effect
for women would suggest no (Calogero, 2004; Gay \& Castano, 2010;
Gervais et al., 2011; Saguy et al., 2010; Yilmaz \& Bozo, 2019), but,
there is evidence that women do objectify other women (Harsey \&
Zurbriggen, 2020; Loughnan et al., 2015a; Puvia \& Vaes, 2013) and to
would be helpful to understand the consequencs of this intragroup
objectification for women. When women objectify other women, does it
lead to self-objectification in the same way men's objectification of
women does (Garcia et al., 2016)? If so, does the self-objectification
experienced in these same-gender interactions have the same negative
consequences for authenticity in that interaction?

@@

In the current study, we sought to examine what occurs during an
interaction in which one or both partners are objectifying each other,
similarly to Garcia et al. (2016), but between same-gender interpersonal
interactions among women. Moreover, the current study uses a
face-to-face interaction paradigm and dyadic data analysis techniques to
examine the effects for both women simultaneously. Although the
literature on woman-woman objectification is small, the results are
mixed, and does not cover interpersonal encourters, we expected to
replicate some of the results found in Garcia et al. (2016). Most
importantly, we predicted that being objectified by one's interaction
partner would lead to self-objectification. We also expected that TSO
would moderate this relationship, amplifying the positive assocaition
between being objectified and SSO for women higher in TSO. Here we would
like to note that Puvia and Vaes (2013) would alternatively predict that
women's tendency to self-objectify (TSO) leads them to objectify other
women (more precisely, to dehumanize a sexualized woman). In addition,
Strelan and Hargreaves (2005) and Harsey and Zurbriggen (2020) would
predict the TSO to other-objectification link, further theorizing that
this relationships is mediated by the woman's own state
self-objectification (SSO). This limited, but extant, set of literature
on woman-woman objectification points to a possible alternative model
from the model hypothesized in the current study. Where appropriate, we
report results considering the causal directions implied by this
alternative mediation model.

We also tested whether SSO would, in turn, lead to feelings of
inauthenticity. The investigation of the SSO - inauthenticity connection
is considered exploratory given the lack of support for this connection
from the literature on identity threat in intragroup interactions
(Rollero, 2016) but support for this connection in the objectification
literature (Garcia et al., 2016; Terán et al., 2020). Regardless of
whether there was an association between SSO and inauthenticity, we
hypothesized that feelings of inauthencity would be associated with
reduced feelings of agency in romantic relationships, reduced career
aspiration, and reduced cognitive performance. In summary, we expected
to find a positive relationship between other-objectification by one's
partner and state self-objectification. We also expected to find a
negative relationship between self-state objectification and interaction
authenticity, and that interaction authenticity will be positively
related to cognitive performance, relationship agency, and career
aspirations.

@@

Thirty-two previously unacquainted dyads of self-identifying women
participated in this study. Data from two demographically similar higher
education institutions in the Northeast United States were combined to
create the final analysis sample (\emph{N =} 64) used in this study. In
the measures section that follows we refer to them as Sample 1 and
Sample 2. Sample 1 (\emph{N =} 24) is from a co-ed liberal arts college
and Sample 2 (\emph{N =} 40) is from a women's liberal arts college.
More specifically, twelve of the pairs were students at a co-ed liberal
arts college, while the remaining twenty pairs attended a women's
liberal arts college.\footnote{Initially, data was collected from both
  same-gender and mixed-gender dyads at both institutions. Sample 1
  originally consisted of 22 pairs, 12 men and 32 women. In Sample 2
  there were 23 pairs made up of 43 women and one man, as well as two
  participants who did not identify with either gender category. To
  investigate only same-gender pairs of women, we limited participant
  data to women in both samples.} All results presented below are from
models including sample as a control variable.

The participants were mostly first-year college students, with an
average age of 18.85 (SD = 1.04). The sample was 48.44\% White/European
American, 9.38\% Black/African-American, 28.12\% Asian/Pacific Islander,
9.38\% Latinx, and 4.69\% mixed-race. There were 8 White/White pairs and
4 same race racial minority pairs, for a total of 12 same-race pairs.
The remaining 20 were mixed race pairs, of which 15 were White/racial
minority pairings and 5 were cross-racial minority group pairs. 64.06\%
of the sample identified as heterosexual, and 25\% identified as gay,
lesbian or bisexual. See the analysis of differences between
heteroseuxal and non-heterosexual women in levels of
other-objectification of their female partners in the measures section
below.

@@

To conceptualize participants' career aspirations after the interaction,
we used the 10-item adaptation of P. Gray and M. OBrien (2007)'s Career
Aspiration Scale which asked participants to consider how true 10
statements were in regard to their future careers on a scale from 0 (not
at all true of me) to 4 (very true of me). Items include \enquote{I hope
to become a leader in my career field} and \enquote{I hope to move up
through any organization or business I work in.} Items were fairly
reliable (\(\alpha\) = 0.73 for Sample 1, \(\alpha\) = 0.80 for Sample
2).

@@

Trait self-objectification (TSO) was assessed using the
Self-Objectification Questionnaire (Barbara L Fredrickson et al., 1998;
M. Noll \& L. Fredrickson, 1998), which evaluates the extent to which
individuals view their bodies in observable versus non-observable ways.
The questionnaire asked participants to rank order both appearance and
functional aspects of their bodies, from 1 (least important) to 10 (most
important), with respect to physical self-concepts. Of the ten body
attributes, five of the items were appearance-based (weight, sex appeal,
physical attractiveness, firm/sculpted muscles and body measurements),
and five of the items were competence-based (strength, physical
coordination, energy level, health and physical fitness). Difference
scores were computed by subtracting the sum of the 5 functional
aspects/competence attributes (e.g., health, strength) from the sum of
the 5 physical self-concepts/appearance attributes (e.g., physical
attractiveness, weight), and all measures were multiplied by -1 so that
positive scores indicated greater TSO.

@@

The R-package \emph{papaja} (Version 0.1.0.9842; Aust \& Barth, 2018)
was used to create a fully reproducible research document containing the
analysis code and manuscript text integrated into one source file. The
benefit of this integration is that the numbers reported throughout this
document are softcoded into the text, ensuring that no errors were made
during a workflow marked by a copy-and-paste from statistical software
to word processing software. Note that here we are using the term
reproducibility to mean getting the same results when running analyses
again using the same data, not to be confused with replicability meaning
collecting a new dataset using the same methods and obtaining the same
results (Patil, Peng, \& Leek, 2016). The source code for this
manuscript along with survey materials and experimenter scripts for the
study protocol are all avaliable at
\url{https://github.com/kkyuchukova/object-in-action}. The data
unfortunutely cannot be made publically avaliable given the dyadic
nature of the observations. For example, with dyadic data, if a person
who participated in the study found their own scores, using the dyad
identification number, they could then see their partner's scores and
thus confidentiality would be broken.

@@

This study sought to detect whether partners' objectification of one
another affected state self-objectification (SSO). Specifically, we were
interested in testing the relationship between state-other
objectification and SSO, and how SSO in turn, may or may not affects
feelings of inauthenticity during the interaction. In addition, we also
tested if the effect of other-objectification in an interaction on SSO
is only present for those women who are high in trait
self-objectification (moderation effect). Further, we investigated the
relationships between experiencing interaction inauthenticity and
relationship agency, career aspirations, and cognitive performance.

While past studies investigating objectification in interactions used
dyadic path analysis (Garcia et al., 2016), the current study used
multilevel modeling proceedures. Dyadic analyses for distinguishable
dyads (e.g., mixed-gender interacting pairs) is more natural in
Structural Equation Modeling (SEM) than it is for indistinguishable
dyads (e.g., same-gender interacting pairs), as we had in the current
study (Garcia, Kenny, \& Ledermann, 2015; Ledermann \& Kenny, 2017). One
reason for this asymmetry is that, due to the arbitrary distinctions
made between \enquote{partner 1} and \enquote{partner 2} in
indistinguishable dyads, many estimates need to be fixed to be equal
(i.e., paths, variances, covariances, endogenous intercepts, and
exogenous means), but these equality constraints should not then be
considered in the degrees of freedom calculations for fit estimations
(Olsen \& Kenny, 2006). Further, Olsen and Kenny (2006) details how a
new independence model and the corresponding fit measure should be
re-calculated for indistinguishable dyads models. The current study used
dyadic multilevel modeling (MLM) to test all relationships, moderation,
and mediation patterns. The online supplementary materials found at
\url{https://github.com/kkyuchukova/object-in-action} contain model
estimates obtained with SEM. See Ledermann and Kenny (2017) for a more
complete discussion of the considerations for using SEM versus MLM for
dyadic analysis.

Testing hypotheses and exploring relationships in the current sample of
indistinguishable dyads involved using the Actor-Partner Independence
Model (APIM) approach for each outcome variable. The current study deals
with indistinguishable dyads, meaning the designation of who is
designated as \enquote{actor} and who is designated as \enquote{partner}
is arbitrary. Thus, we ran five APIM's to test all the relationships
(i.e., for SSO, inauthenticity, career aspirations, relationship agency,
and cognitive performance as outcome variables). See
Figure~\ref{fig:apim} for a basic APIM model. The APIM includes effects
due to one's own, as well as one's partner's, predictor variables
(\(X\)'s) on the one's own outcome variable (\(Y\)). The analysis of the
relationship between SSO and inauthenticity in this same-gender sample
are considered exploratory, given that prior research points to mixed
possibilities.

@@

\begin{figure}
\includegraphics[width=400px]{SEMfigure} \caption{This figure depicts the relationships tested bewteen study variables. The light gray variables represent redundant variables and are shown to emphasize the dyadic nature of the data. The estimates were from obtained from separate MLMs. Sample was controlled for in all models. The effects of TSO, actor other-objectification, and partner other-objectification are from a model where the interaction of TSO and partner other-objectification had been trimmed. *p < .05, dashed lines represent non-significant relationships.}\label{fig:semfigure}
\end{figure}

All relationships between study variables and the MLM estimates are
depicted in Figure~\ref{fig:semfigure}.

@@

As is the case in all correlational studies, we cannot be sure about the
causal direction. The model used by the current study asserts that being
objectified by a woman causes women to self-objectify, but it could be
that women's SSO causes them to be objectified by their interaction
partner. This latter interpretation is theoretically interesting given
that, in both the Garcia et al. (2016) study and in the current study,
it is \emph{women's} state self-objectification that relates to being
objectified by one's partner. Objectification Theory (Barbara L
Fredrickson \& Roberts, 1997), as well as some past experimental studies
(Saguy et al., 2010) suggest that at least one direction of the causal
flow is from other-objectification to SSO. Other studies, especailly
those investigating explicitly women objectifying other women (Puvia \&
Vaes, 2013), have relately found evidence that women's \emph{trait}
self-objectification causes women to objectify (dehumanize) other women
and this link is mediated by \emph{state} self-objectification. It could
follow that we tend to objectify other women who objectify us and
through a process of TSO causes SSO which in turn causes objectifation
of one's partner, which in turn causes one's partner to objectify us.
The current study does not provide any evidence of this alternative
model, however---TSO was not significantly correlated with SSO and
actor's objectification was not correlated with partner's
objectification.

Where the results of the current study diverge most notably from the
results of studies testing interpersonal objectification among
mixed-gender dyads is the lack of evidence for relationships between SSO
and interaction inauthenticity. In addition, we found no evidence of
\emph{direct} or \emph{indirect} relationships between SSO and cognitive
functioning, relationship agency, and career aspirations. Nor was there
evidence of \emph{direct} or \emph{indirect} relationships between
other-objectification and any of these outcomes. This is somewhat
suprising given the extant evidence linking SSO and cognitive
functioning (Barbara L Fredrickson et al., 1998; and see Moradi \&
Huang, 2008 for a review; Quinn, Chaudoir, \& Kallen, 2011) and the
research on interpersonal other-ojectification and cognitive functioning
(Garcia et al., 2016; Gervais et al., 2011; Logel et al., 2009). This
lack of evidence could potentially signal diverging processes between
women's experiences with interpersonal objectification from men and
interpersonal objectification from women. There is quite a bit of
evidence suggesting that the male gaze is particularly detrimental
(Calogero, 2004; Barbara L Fredrickson \& Roberts, 1997; Gervais,
Holland, \& Dodd, 2013; Gervais et al., 2011), and perhaps the
self-objectification experienced within an interaction with a women is
qualtatively different, and perhaps not as harmful, as the
self-objectification experienced within an interaction with a man.
However, as a strong note of caution, it is best not to interpret a null
result as evidence of no relationship and more research on interpersonal
objectification among women is needed.
